---
title: "R Skills - data2pdf"
author: "Bill Foote"
date: "January 4, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
The objective is to show the work flow from data collection through output and rendering.

## Data

Data is from [FRED](https://fred.stlouisfed.org/categories/32407).

\begin{center}
  \begin{tabular}{| l | p{9cm} | l |}
    \hline
    Variable & Description & Units \\ \hline
    rate & Weighted-Average Effective Loan Rate by Time that Pricing Terms Were Set and by Commitment, Before Survey Week, More than 365 Days, All Commercial Banks (EETP365PDNQ) & percent \\ \hline
prepay & Percent of Value of Loans Subject to Prepayment Penalty by Time that Pricing Terms Were Set and by Commitment, Before Survey Week, More than 365 Days, All Commercial Banks (EPTP365PDNQ) & percent \\ \hline
maturity & Weighted-Average Maturity/Repricing Interval in Days by Time that Pricing Terms Were Set and by Commitment, Before Survey Week, More than 365 Days, All Commercial Banks (EITP365PDNQ) & days \\ \hline
size & Average Loan Size by Time that Pricing Terms Were Set and by Commitment, Before Survey Week, More than 365 Days, All Commercial Banks (EATP365PDNQ) & thousandsUSD \\ \hline
volume & Total Value of Loans by Time that Pricing Terms Were Set and by Commitment, Before Survey Week, More than 365 Days, All Commercial Banks (EVTP365PDNQ) & millionsUSD \\ \hline
  \end{tabular}
\end{center}

## Data Exploration

Read in the data.

```{r}
x.data <- read.csv("data/commloans.csv")
head(x.data)
tail(x.data)
summary(x.data)
```

## Analysis

Insert analysis here.


## Observations & Recommendations

Insert observations and recommendations summary here.